---
title: "purrr-RLadies-talk"
author: "Phoebe Wong"
date: "January 21, 2018"
output: 
  md_document:
  variant: markdown_github
---

The idea of using purrr is to solve iteration problems by using:

1. functions that write for loops for you
2. with consistent syntax & output, and
3. convenience shortcuts for specifying functions to iterate 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

## Setup
Checking if you have the packages
```{r, message=FALSE}
library(tidyverse)
library(repurrrsive) #devtools::install_github("jennybc/repurrrsive")
```

We will working with the sw_people dataset that is in repurrrsive package. It's a dataset about Star Wars characters, queried from the [Star Wars API](https://swapi.co/).

Now, let's get our hands dirty and look at the data.

First, let's examine sw_people.

```{r}
class(sw_people)
```

Okay, now we know that sw_people is a list. Let's see how many elements are there in sw_people.
```{r}
length(sw_people)
```

YOUR TURN:

1. Who is the first person listed in sw_people? 
2. What information is given for the first person?
3. What is the difference between sw_people[1] and sw_people [[1]]

Wow, there is quite a lot of information for Luke and there're more than one character in Star Wars! 
I'm interested to see how many starships has each character been in.

First, let's start small by finding out how many starships does **Luke** have.
```{r}
length(sw_people[[1]]$starships)
```
So, if I want to know how many starships other characters have, what should I change?
```{r, eval=FALSE}
length(sw_people[[2]]$starships)
length(sw_people[[3]]$starships)
length(sw_people[[4]]$starships)
```

But there are so many characters in sw_people, how can we do that more systemtically? 

# map()!

`map()` helps to solve an iteration problem like "for each element of .x for .f" by using **map(.x, .f, ...)**.

In ``map()``, there are two important inputs.

* .x: the element you want to loop through: a vector, a list or a data frame (for each column)
* .f: a function that you want to be applied to each element

In our example above, we want to solve: for **each person** in sw_people, the **number of starships** they have. As noted above, we can use `length(x$starships)` to get the count of number of starships they have.

```{r}
map(sw_people, ~length(.x$starships)) %>% head()
```

For those who are familiar with for loops or s/lapply(), you will know that we can get the same numbers by doing a foor loop:
```{r}
num_starship <- numeric()
for (i in 1:length(sw_people)){
  num_starship[i] <- length(sw_people[[i]]$starships)
}
num_starship
```
or a sapply()
```{r}
sapply(sw_people, function(x) length(x$starships))
```

# Why should we use `map()`?
The main advantage of `map()` is the helpers which allow you to write compact code for common special cases. It also has much greater consistency between the functions, for example, the first argument to all map functions is always the data while it's different between `lapply()` and `mapply()`. Hadley also commented on the comparison of purrr and lapply on this [Stack Overflow post](https://stackoverflow.com/questions/45101045/why-use-purrrmap-instead-of-lapply/47123420#47123420) and his [Twitter post](https://twitter.com/hadleywickham/status/927199362668814336).

Let's try to use map for another example. Let's find the name of each character's home world. 

We will use the variable, planet_lookup that is created using the following code, but we can ignore the code details for now.

```{r}
planet_lookup <- map_chr(sw_planets, "name")  %>%  # planets 
  set_names(map_chr(sw_planets, "url")) #name as the URL
head(planet_lookup)
```

planet_lookup is a named character that the URLs are the names and the home world planet names are the value. In our sw_people data, the home world of each character is stored in sw_people[[i]]$homeworld as a URL. 
```{r}
sw_people[[1]]$homeworld
```

We can use the URL as an index to find the planet name of Luke from planet_lookup.
```{r}
planet_lookup[sw_people[[1]]$homeworld]
```

Your turn! Can you use `map()` to find the name of each character's home world?

**Tips**: think about what do we need to change to find it for each character

# Other map functions
As you can notice from our above example, the output from `map()` is always a **list**. 

Sometimes, we want to have the output in other simpler formats. There are a few other functions in purrr that can help:

1. **map_lgl()** returns logical vector
2. **map_int()** returns integer vector
3. **map_dbl()** returns double vector (What is [double vector]?(http://uc-r.github.io/integer_double/))
4. **map_chr()** returns a character vector

Another function, `walk()` returns nothing at all, it's helpful in cases where you:

1. only want to print to screen
2. plot to graphics device
3. file maniuplation (saving, writing, moving, etc.)
4. system calls

Your turn!
Can you use map and the other typed functions to answer the following question?

1. How many starships has each character been in?
2. What color is each character's hair?
3. Is the character male?
4. How heavy is each character?

To find out how heavy is each character, we want the **numeric** value of "mass" of each character from the dataset. 

If we use `map_dbl()`, it will not work, because "mass" is stored as a **string** in the data

```{r}
map_dbl(sw_people, ~.x[["mass"]])
```

```{r}
map(sw_people, ~ .x[["mass"]]) %>% head()
```
The solution is to do it in two steps.

```{r}
map_chr(sw_people, ~ .x[["mass"]]) %>% # map the character
  readr::parse_number(na = "unknown") #convert character to numbers and turn "unknown" to NA
```

# .f can be a formula, a string or an integer
```{r, results="hide"}
map_chr(sw_people, ~ .x[["hair_color"]]) # a formula
# can be
map_chr(sw_people, "hair_color") # a string
```

```{r, results="hide"}
char_starships <- map(sw_people, "starships")
map_int(char_starships, length) # a function
# equivalent to
map_int(sw_people, ~length(.x[["starships"]])) # a formula
```

Your turn!
Star Wars Challenges

1. Which film (sw_films) has the most characters?
2. Which sw_species has the most possible eye colors?
3. Which sw_planets do we know the least about (i.e., have the most "unknown" entries)?

# What if I have more than 1 variable to iterate?
## map2()!

Similar to `map()`, `map2()` helps with iteration problems with two variables.
map2(.x, .y, .f): for each element of .x and corresponding element of .y, apply .f
```{r}
letter <- c("a", "b", "c")
number <- c(1, 2, 3)
map2(letter, number, paste)
```

It can be helpful for plotting.
```{r}
gap_split_small <- gap_split[1:10]
countries <- names(gap_split_small)

# For each country create a ggplot of Life Expectancy through time 
# with a title

# For one country
ggplot(gap_split_small[[1]], aes(year, lifeExp)) +
  geom_line() +
  labs(title = countries[[1]])

# For all countries
plots <- map2(gap_split_small, countries, 
  ~ ggplot(.x, aes(year, lifeExp)) + 
      geom_line() +
      labs(title = .y))

plots[[1]]
# Display all plots
walk(plots, print) # this might take awhile

# Save all plots
walk2(.x = plots, .y = countries, 
  ~ ggsave(filename = paste0(.y, ".pdf"), plot = .x))

# Argh! I didn't want all those pictures in this directory,
# remove them all
file.remove(paste0(countries, ".pdf"))

```

Similar to `map()`, there are also other type functions for `map2()`: `walk2()`, `map2_lgl()`, `map2_int()`, `map2_dbl()`, `map2_chr()` that will give you outputs in other formats.

# What if I have more than 2 variables to iterate?
## pmap()!
pmap(.l, .f, ...): for each element of each vector in .l, apply .f

# What if I want to apply multiple functions on the same variable?
## invoke_map()!
invoke_map(.f, .x, ...): for each function in .f, apply to .x

# Reference
Most of the material references from Charlotte Wickham's [purrr tutorial](https://github.com/cwickham/purrr-tutorial). There are also some [challenges](https://github.com/cwickham/purrr-tutorial/tree/master/challenges) on the repo that you should try!


